### 2019-05-09 initial/notes.tex
okay get Phop correctly

\section{Hamiltonians and EOM}

@@

\subsection{Try 1: Qualitative}

We zoom in on Cassini State $4$, which has $\theta_4 =
-\frac{\pi}{2} + \frac{\eta \cos I}{1 - \eta \sin I}, \mu_4 = \frac{\eta \cos
I}{1 - \eta \sin I}, \phi_4 = 0$. Then, using
equations of motion
\begin{align}
    \pd{\phi}{t} &= \mu - \eta\p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}}
        \cos \phi} ,\\
    \pd{\mu}{t} &= -\eta \sin I \sin \phi + \s*{\epsilon \p*{1 - \mu^2}},
\end{align}
we can perturbatively require $\pd{\theta}{t} = 0$ for $\epsilon \neq 0$. This
corresponds to $\eta \sin I \sin \p*{\phi_4 + \delta \phi} \approx \epsilon$, or
$\delta \phi_4 = +\frac{\epsilon}{\eta \sin I}$. This is in agreement with
Dong's result. Note that $\delta \theta_2 = -\frac{\epsilon}{\eta \sin I}$,
which I saw in my simulations.

This implies that the stable manifolds of the two saddle points, which once
overlapped with each other's unstable manifolds (creating a heteroclinic orbit)
now are offset from one another by distance $D \sim \frac{\epsilon}{\eta \sin
I}$. The question is how likely it is to thread the needle.

Consider that, very near CS4, the angle of incidence on the desired gap is
roughly $\tan \psi \approx \psi = \frac{\Delta \theta}{\Delta \phi}$. Over the
course of one orbit, $\Delta \phi$ changes by $2\pi$, while $\Delta \theta \sim
\epsilon \sin \theta T$ where $T$ is the period of an orbit. Examining the data,
$T \sim 50$, and so $\frac{\Delta \theta}{\Delta \phi} \sim \frac{2\pi}{\epsilon
(50)}$.

The effective probability of threading the opened gap between the
stable/unstable manifolds is then just $P \propto D\sin \psi \sim
\frac{2\pi}{T\p*{\eta \sin I}}$. According to later analysis, this should really
be $\frac{2\pi}{T}$. Plugging in some observational values $T \sim 50$ for $\eta
= 0.1$gives $P \propto 0.13$. In reality, I find it asymptotes to $\sim 0.08$,
so the constant of proportionality is of order unity. Not bad given the really
crappy $\psi \sim \frac{\ev*{\dot{\theta}}}{\dot{\phi}}$ argument.

\subsection{Try 2: Melnikov Distance}

@@

Let's apply this to the Cassini state Hamiltonian w/ dissipation. We first write
down our EOM in Melnikov form (we use canonical variables $\mu, \phi$):
\begin{align}
    \rd{\hat{s}}{t}
        &= \underbrace{
            \pd{\mathcal{H}}{\mu}\hat{\phi} - \pd{\mathcal{H}}{\phi}\hat{\mu}}_f
            + \epsilon \underbrace{\p*{1 - \mu^2}\hat{\mu}}_g.
\end{align}
Then $f \times g = f_\phi g_\mu = \pd{\mathcal{H}}{\mu}\p*{1 - \mu^2}$. We then
want to integrate this along the heteroclinic orbit. We can make change of
variables
\begin{equation}
    M = \int\limits_0^{2\pi}\pd{\mathcal{H}}{\mu}
        \p*{1 - \mu^2}\p*{\pd{\phi}{t}}^{-1}\;\mathrm{d}\phi.
\end{equation}
But thankfully, $\pd{\mathcal{H}}{\mu} = \pd{\phi}{t}$ in the absence of
dissipation, and so $M = 2\pi\p*{1 - \mu^2} \approx 2\pi\p*{1 - 2\eta \sin I}$.
Thus, the Melnikov distance at point $q^0$, a point on the heteroclinic
orbit of the unperturbed Hamiltonian, is just
\begin{equation}
    d(q^0) = \frac{2\pi \epsilon\p*{1 - 2\eta \sin I}}{\abs*{f(q^0)}}.
\end{equation}
Note that the maximum value $\abs*{f(q^0)}$, which occurs at $\phi = \pi$, is
just $f \approx \sqrt{4\eta \sin I}$.

@@

\subsection{Try 3: Poincar\'e Section}

@@

Abandoning the rest of this for the time being.

>>>>

\section{Constant $\eta$}

@@

\subsection{Consideration 1: Qualitative}

We zoom in on Cassini State $4$, which has $\theta_4 =
-\frac{\pi}{2} + \frac{\eta \cos I}{1 - \eta \sin I}, \mu_4 = \frac{\eta \cos
I}{1 - \eta \sin I}, \phi_4 = 0$. Then, using
equations of motion
\begin{align}
    \pd{\phi}{t} &= \mu - \eta\p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}}
        \cos \phi} ,\\
    \pd{\mu}{t} &= -\eta \sin I \sin \phi + \s*{\epsilon \p*{1 - \mu^2}},
\end{align}
we can perturbatively require $\pd{\theta}{t} = 0$ for $\epsilon \neq 0$. This
corresponds to $\eta \sin I \sin \p*{\phi_4 + \delta \phi} \approx \epsilon$, or
$\delta \phi_4 = +\frac{\epsilon}{\eta \sin I}$. This is in agreement with
Dong's result.

This implies that the stable manifolds of the two saddle points, which once
overlapped with each other's unstable manifolds (creating a heteroclinic orbit)
now are offset from one another by distance $D \sim \frac{\epsilon}{\eta \sin
I}$. But since $\epsilon$ also sets $\dot{\mu}$ in a precession orbit-averaged
sense, the effective cross section is constant in some sense: there will be one
orbit where $\mu$ goes from below CS4 to above CS4, during which it will make
jump of size $\epsilon$, and if it hits a particular interval of size $\epsilon$
then it will enter the separatrix. Thus, separatrix hopping should $\propto
\epsilon^0$.

\subsection{Consideration 2: Melnikov Distance}

@@

Let's apply this to the Cassini state Hamiltonian w/ dissipation. We first write
down our EOM in Melnikov form (we use canonical variables $\mu, \phi$):
\begin{align}
    \rd{\hat{s}}{t}
        &= \underbrace{
            \pd{\mathcal{H}}{\mu}\hat{\phi} - \pd{\mathcal{H}}{\phi}\hat{\mu}}_f
            + \epsilon \underbrace{\p*{1 - \mu^2}\hat{\mu}}_g.
\end{align}
Then $f \times g = f_\phi g_\mu = \pd{\mathcal{H}}{\mu}\p*{1 - \mu^2}$. We then
want to integrate this along the heteroclinic orbit. We can make change of
variables
\begin{equation}
    M = \int\limits_0^{2\pi}\pd{\mathcal{H}}{\mu}
        \p*{1 - \mu^2}\p*{\pd{\phi}{t}}^{-1}\;\mathrm{d}\phi.
\end{equation}
But thankfully, $\pd{\mathcal{H}}{\mu} = \pd{\phi}{t}$ in the absence of
dissipation, and so $M = 2\pi\p*{1 - \mu^2} \approx 2\pi$.
Thus, the Melnikov distance at point $q^0$, a point on the heteroclinic
orbit of the unperturbed Hamiltonian, is just
\begin{equation}
    d(q^0) = \frac{2\pi \epsilon}{\abs*{f(q^0)}}.
\end{equation}
Note that the maximum value $\abs*{f(q^0)}$, which occurs at $\phi = \pi$, is
just $f_{\max} \approx \sqrt{4\eta \sin I}$.

@@

\subsection{Consideration 3: Poincar\'e Section}

@@

\subsection{Consideration 4: Plotting Stable/Unstable manifolds}

We can plot the stable/unstable manifolds of two Cassini States as in
\autoref{fig:CS4s}. Then, since phase space is roughly flat near $\phi = \pi$
(near $\phi = 0$, $\dot{\phi}$ varies drastically and so phase space is
``squished'' a bit via Liouville's Theorem), we just need to compare the
distance between $\mathcal{W}^{(0)}_S$ and $W_U^{(0)}$, the capture gap, to the
distance between $\mathcal{W}^{(1)}_S$ and $W_U^{(0)}$ the Melnikov gap, to
estimate the capture probability.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.6\textwidth]{0_eta/6manifolds0_10.png}
    \caption{Stable/Unstable manifolds of the two Cassini State
    4s.}\label{fig:CS4s}
\end{figure}

The Melnikov gap is predicted above as $d(q^0)$ or approximately
\begin{equation}
    \Delta_{M} \approx \frac{2\pi \epsilon}{\sqrt{4\eta\sin I}}.
\end{equation}

The capture gap is much trickier to predict, since it depends on the separation
between $\mathcal{W}^{(0)}_S$ \emph{after passing through} CS$_4^{(1)}$.
Instead, let's consider the closed orbit in the absence of dissipation that
starts at CS4', the modified CS\@. This orbit has a finite period set by
equating $\int\limits_0^T \pd{\phi}{t}\;\mathrm{d}t =
\int\limits_0^{2\pi}\;\mathrm{d}\phi + \int\limits_{2\pi}^0\;\mathrm{d}\phi$.

Now, let's reconsider the Melnikov integral when perturbing this finite
(non-homoclinic orbit); this may no longer be an exact result but should give
the correct scaling:
\begin{equation}
    M_c = \int\limits_0^T \pd{\phi}{t}\epsilon \p*{1 - \mu^2}\;\mathrm{d}t.
\end{equation}
Naively, we might claim that, since $\pd{\phi}{t}$ changes signs halfway through
the interval of integration, that the only surviving component is $2\epsilon
\bar{\mu}\mu'$, where $\bar{\mu} = \mu_4$ is the average value of $\mu$ and
$\mu'$ is the fluctuation. This gives
\begin{equation}
    M_c = 2\int\limits_0^{2\pi}2\epsilon \frac{\eta \cos I}{1 + \eta \sin I}
        \sqrt{2\eta \sin I \p*{1 - \cos \phi}}\;\mathrm{d}\phi.
\end{equation}
Note that $M_c \propto \eta^{3/2}$, and since the gap opened $\Delta_c =
\frac{M_c}{\pd{\phi}{t}} \propto \eta$, it seems like we're on the right track.
Specifically:
\begin{align}
    M_c &\approx \epsilon 2\eta \cos I A_{sep},\\
    \Delta_c &\approx 2\epsilon \eta \cos I \p*{16 \sqrt{\eta \sin I}}
        \frac{1}{\sqrt{4\eta \sin I}},\\
        &\approx 8\epsilon \eta \cos I.
\end{align}
This also agrees exceedingly well with our simulations This then gives us
hopping probability
\begin{equation}
    P_{hop} = \frac{\Delta_c}{\Delta_M} \approx
        \frac{16\eta^{3/2}\cos I \sqrt{\sin I}}{\pi}.
\end{equation}
This agrees perfectly with the cases we've run.
